Author disambiguation of large-scale semi-labeled collections of scientific articles

### 2015-08-19 paper.tex
supervised learning

% For placing tables in correct place
\usepackage{placeins}

@@

The remaining of this report is structured as follows. In Section~\ref{related-works},
we first briefly review machine learning solutions for author disambiguation.
The components of our method are then defined in Section~\ref{methods}
and its implementation described in Section~\ref{implementation}. Experiments
are carried out in Section~\ref{experiments}, where we explore and validate
features for the supervised learning of a similarity function and compare
strategies for the semi-supervised clustering of publications.
Finally, conclusions and future works are discussed in Section~\ref{conclusions}.

@@

As reviewed in
\citep{smalheiser2009author,ferreira2012brief,levin2012citation}, author
disambiguation algorithms are usually composed of two main components: (i) a
similarity function determining whether two publications have been written by a
same author and (ii) a clustering algorithm producing clusters of publications
assumed to be written by a same author. Approaches can be inventorized along
several axes, depending on the type and amount of data available, the way the
similarity function is learned or defined, or the clustering procedure used to
group publications. Methods relying on supervised learning usually make use of
a small set of
hand-labeled pairs of publications identified as being either from the same or
from different authors for learning automatically a similarity function between
publications \citep{han2004two,huang2006efficient,
culotta2007author,treeratpituk2009disambiguating,tran2014author}. Because
training data is often not easily available, unsupervised approaches propose
instead to use domain-specific similarity functions tailored for author
disambiguation \citep{malin2005unsupervised,mcrae2006also,song2007efficient,
soler2007separating, kang2009co,fan2011graph,schulz2014exploiting}. These later
approaches have the advantage of not requiring  hand-labeled data, but often do
not perform as well as the supervised approaches. To reconcile both worlds,
semi-supervised methods make use of small, manually verified, clusters of
publications and/or of high-precision domain-specific rules to build a training
set of pairs of publications, from which a similarity function is then built
using supervised learning
\citep{ferreira2010effective,torvik2009author,levin2012citation}.
Coincidentally, semi-supervised approaches also allow for the tuning of the
clustering algorithm when the later is applied to a mixed set of labeled
and unlabeled publications, e.g., by maximizing some clustering performance
metric on the known clusters \citep{levin2012citation}.

@@

In this context, we position the work presented in this paper as a
semi-supervised  solution for author disambiguation, but with the significant
advantage of having a very large collection of more than 1 million crowdsourced annotations
of publications whose true authors are identified. In particular, the extent and coverage
of this data allows us to revisit, validate and nuance previous findings regarding
supervised learning of similarity functions and to better explore strategies
for semi-supervised clustering. Additionally, by releasing the data publicly,
we hope to set a benchmark on which further research on author disambiguation
and related topics can  build upon.

@@

As inspired from several previous works described in Section~\ref{related-works},
we cast in this work author disambiguation into a semi-supervised clustering
problem.  Our algorithm is made of three parts: (i) a blocking
scheme whose goal is to rhoughly pre-cluster signatures ${\cal S}$ into smaller groups in order to
reduce computational complexity; (ii) the construction of a similarity function
$d$ between signatures using supervised learning; (iii) the
semi-supervised clustering of all signatures within a same block, using $d$ as a pseudo distance metric.

@@

\subsection{Learning a similarity function}
\label{methods:similarity}
\glnote{By Gilles}

@@

As part of this work, we developed a stand-alone application for author
disambiguation, publicly available
online\footnote{\url{https://github.com/inveniosoftware/beard}} for free reuse
or study.  Our implementation builds upon the Python scientific stack, making
use of the Scikit-Learn library \citep{scikitlearn} for the supervised learning
of a similarity function and of SciPy \citep{scipy} for clustering. All
components of the disambiguation pipeline have been designed to follow the
Scikit-Learn API \citep{scikitlearnAPI}, making them easy to maintain,
understand and reuse.

@@

\FloatBarrier
\begin{table}[H]
\caption{Base experiment evaluation)}
\centering
\begin{tabular}{|r|r|l|l|l|}
  \hline
  Evaluation & Set & Precision & Recall & F-score \\
  \hline
  Pairwise & test & 0.9948 & 0.9738 & 0.9842 \\
  \hline
  Pairwise & train & 0.9976 & 0.9747 & 0.986 \\
  \hline
  Pairwise & overall & 0.9952 & 0.9739 & 0.9844 \\
  \hline
  $B^3$ & test & 0.9901 & 0.976 & 0.983 \\
  \hline
  $B^3$ & train & 0.997 & 0.979 & 0.9879 \\
  \hline
  $B^3$ & overall & 0.9907 & 0.9761 & 0.9834 \\
  \hline
\end{tabular}
\end{table}

@@

\FloatBarrier
\begin{table}[H]
\caption{Blocking limited maximum recalls (BLMR) on the whole dataset without splitting over
the first given name initial}
\centering
\begin{tabular}{|r|l|l|}
  \hline
  blocking function & BLMR with $B^{3}$ recall & BLMR with pairwise recall \\
  \hline
  last name and first initial & 0.9827 & 0.9776 \\
  \hline
  phonetic using double metaphone & 0.9967 & 0.9953 \\
  \hline
  phonetic using nysiis & 0.9961 & 0.9951 \\
  \hline
  phonetic using soundex & 0.9966 & 0.9953 \\
  \hline
\end{tabular}
\end{table}

@@

\FloatBarrier
\begin{table}[H]
\caption{Phonetic algorithm performance}
\centering
\begin{tabular}{|r|r|r|l|l|l|}
  \hline
  Phonetic Algorithm & Split over initial & Evaluation & Precision & Recall & F-score \\
  \hline
  Double Metaphone & yes & pairwise & 0.9934 & 0.982 & 0.9876 \\
  \hline
  Double Metaphone & yes & $B^3$ & 0.9867 & 0.983 & 0.9848 \\
  \hline
  Double Metaphone & no & pairwise & - & - & - \\
  \hline
  Double Metaphone & no & $B^3$ & - & - & - \\
  \hline
  NYSIIS & yes & pairwise & 0.9942 & 0.9818 & 0.9879 \\
  \hline
  NYSIIS & yes & $B^3$ & 0.9884 & 0.9828 & 0.9856 \\
  \hline
  NYSIIS & no & pairwise & - & - & - \\
  \hline
  NYSIIS & no & $B^3$ & - & - & - \\
  \hline
  Soundex & yes & pairwise & - & - & - \\
  \hline
  Soundex & yes & $B^3$ & - & - & - \\
  \hline
  Soundex & no & pairwise & - & - & - \\
  \hline
  Soundex & no & $B^3$ & - & - & - \\
  \hline
\end{tabular}
\end{table}

@@

\subsection{Learning a similarity function}
% feature engineering
% variable importance analysis
% supervised learning model

>>>>

@@

The remaining of this report is structured as follows. In Section~\ref{related-works},
we first briefly review machine learning solutions for author disambiguation.
The components of our method are then defined in Section~\ref{methods}
and its implementation described in Section~\ref{implementation}. Experiments
are carried out in Section~\ref{experiments}, where we explore and validate
features for the supervised learning of a linkage function and compare
strategies for the semi-supervised clustering of publications.
Finally, conclusions and future works are discussed in Section~\ref{conclusions}.

@@

As reviewed in
\citep{smalheiser2009author,ferreira2012brief,levin2012citation}, author
disambiguation algorithms are usually composed of two main components: (i) a
linkage function determining whether two publications have been written by a
same author and (ii) a clustering algorithm producing clusters of publications
assumed to be written by a same author. Approaches can be inventorized along
several axes, depending on the type and amount of data available, the way the
linkage function is learned or defined, or the clustering procedure used to
group publications. Methods relying on supervised learning usually make use of
a small set of
hand-labeled pairs of publications identified as being either from the same or
from different authors for learning automatically a linkage function between
publications \citep{han2004two,huang2006efficient,
culotta2007author,treeratpituk2009disambiguating,tran2014author}. Because
training data is often not easily available, unsupervised approaches propose
instead to use domain-specific linkage functions tailored for author
disambiguation \citep{malin2005unsupervised,mcrae2006also,song2007efficient,
soler2007separating, kang2009co,fan2011graph,schulz2014exploiting}. These later
approaches have the advantage of not requiring  hand-labeled data, but often do
not perform as well as the supervised approaches. To reconcile both worlds,
semi-supervised methods make use of small, manually verified, clusters of
publications and/or of high-precision domain-specific rules to build a training
set of pairs of publications, from which a linakge function is then built
using supervised learning
\citep{ferreira2010effective,torvik2009author,levin2012citation}.
Coincidentally, semi-supervised approaches also allow for the tuning of the
clustering algorithm when the later is applied to a mixed set of labeled
and unlabeled publications, e.g., by maximizing some clustering performance
metric on the known clusters \citep{levin2012citation}.

@@

In this context, we position the work presented in this paper as a
semi-supervised  solution for author disambiguation, but with the significant
advantage of having a very large collection of more than 1 million crowdsourced annotations
of publications whose true authors are identified. In particular, the extent and coverage
of this data allows us to revisit, validate and nuance previous findings regarding
supervised learning of linkage functions and to better explore strategies
for semi-supervised clustering. Additionally, by releasing the data publicly,
we hope to set a benchmark on which further research on author disambiguation
and related topics can  build upon.

@@

As inspired from several previous works described in Section~\ref{related-works},
we cast in this work author disambiguation into a semi-supervised clustering
problem.  Our algorithm is made of three parts: (i) a blocking
scheme whose goal is to rhoughly pre-cluster signatures ${\cal S}$ into smaller groups in order to
reduce computational complexity; (ii) the construction of a linkage function
$d$ between signatures using supervised learning; (iii) the
semi-supervised clustering of all signatures within a same block, using $d$ as a pseudo distance metric.

@@

\subsection{Linkage function}
\label{methods:linkage}

\textit{Supervised classification.} The second part of the algorithm is the
automatic construction of a pair-wise linkage function between signatures, for
later reuse during clustering to group all signatures from a same author.
Formally, the goal is to build a function $d: {\cal S} \times {\cal S} \mapsto
[0, 1]$, such that $d(s_1, s_2)$ tends to $0$ if both signatures $s_1$ and
$s_2$ belong to the same author, and $1$ otherwise.

As in previous works on author disambiguation, we cast this problem as a
classification task, where inputs are pairs of signatures and outputs are
classes $0$ (same authors) and $1$ (distinct authors).

\textit{Features engineering.}

\textit{Building a training set.}
% cite https://hpi.de/naumann/projects/repeatability/datasets/dblp-dataset.html

@@

As part of this work, we developed a stand-alone application for author
disambiguation, publicly available
online\footnote{\url{https://github.com/inveniosoftware/beard}} for free reuse
or study.  Our implementation builds upon the Python scientific stack, making
use of the Scikit-Learn library \citep{scikitlearn} for the supervised learning
of a linkage function and of SciPy \citep{scipy} for clustering. All
components of the disambiguation pipeline have been designed to follow the
Scikit-Learn API \citep{scikitlearnAPI}, making them easy to maintain,
understand and reuse.

@@

\begin{table}[H]
\caption{Base experiment evaluation)}
\centering
\begin{tabular}{|r|r|l|l|l|}
  \hline
  Evaluation & Set & Precision & Recall & F-score \\
  \hline
  Pairwise & test & 0.9948 & 0.9738 & 0.9842 \\
  \hline
  Pairwise & train & 0.9976 & 0.9747 & 0.986 \\
  \hline
  Pairwise & overall & 0.9952 & 0.9739 & 0.9844 \\
  \hline
  $B^3$ & test & 0.9901 & 0.976 & 0.983 \\
  \hline
  $B^3$ & train & 0.997 & 0.979 & 0.9879 \\
  \hline
  $B^3$ & overall & 0.9907 & 0.9761 & 0.9834 \\
  \hline
\end{tabular}
\end{table}

@@

\begin{table}[H]
\caption{Blocking limited maximum recalls (BLMR) on the whole dataset without splitting over
the first given name initial}
\centering
\begin{tabular}{|r|l|l|}
  \hline
  blocking function & BLMR with $B^{3}$ recall & BLMR with pairwise recall \\
  \hline
  last name and first initial & 0.9827 & 0.9776 \\
  \hline
  phonetic using double metaphone & 0.9967 & 0.9953 \\
  \hline
  phonetic using nysiis & 0.9961 & 0.9951 \\
  \hline
  phonetic using soundex & 0.9966 & 0.9953 \\
  \hline
\end{tabular}
\end{table}

@@

\begin{table}[H]
\caption{Phonetic algorithm performance}
\centering
\begin{tabular}{|r|r|r|l|l|l|}
  \hline
  Phonetic Algorithm & Split over initial & Evaluation & Precision & Recall & F-score \\
  \hline
  Double Metaphone & yes & pairwise & 0.9934 & 0.982 & 0.9876 \\
  \hline
  Double Metaphone & yes & $B^3$ & 0.9867 & 0.983 & 0.9848 \\
  \hline
  Double Metaphone & no & pairwise & - & - & - \\
  \hline
  Double Metaphone & no & $B^3$ & - & - & - \\
  \hline
  NYSIIS & yes & pairwise & 0.9942 & 0.9818 & 0.9879 \\
  \hline
  NYSIIS & yes & $B^3$ & 0.9884 & 0.9828 & 0.9856 \\
  \hline
  NYSIIS & no & pairwise & - & - & - \\
  \hline
  NYSIIS & no & $B^3$ & - & - & - \\
  \hline
  Soundex & yes & pairwise & - & - & - \\
  \hline
  Soundex & yes & $B^3$ & - & - & - \\
  \hline
  Soundex & no & pairwise & - & - & - \\
  \hline
  Soundex & no & $B^3$ & - & - & - \\
  \hline
\end{tabular}
\end{table}

@@

\subsection{Learning a linkage function}
% feature engineering
% variable importance analysis
% supervised learning model
